Future challenges for Research Software Engineering

### 2025-09-19 future_challenges_for_RSEng.tex
Update future_challenges_for_RSEng.tex
Co-authored-by: Jeremy Cohen <jcohen02@users.noreply.github.com>

\subsection{Impact of Alternative Career Paths for Researchers on the Structure of Research Institutes}
\begin{whatis}{}{career-paths}
As the RSE becomes establishedwe can see the emergence of more dedicated staff, i.e. other (digital) RTPs.
The legal requirements for employing people have become flexible enough to allow for the employment of more RTPs.
Work outside of the publication record is more readily accepted and enables researchers to build careers at the interface of academia, industry, and policy
(\eg\ in think tanks, NGOs, or organisations that do knowledge transfer).
\end{whatis}
RSEs can either be embedded in research groups (belonging to the mid-level university staff),
or be members of larger RSE units~\cite{Kempf2025-draft}.
The participants had the opinion that focusing on embedded RSEs would make onboarding new RSEs to existing projects easier.
This, however, comes with additional funding challenges,
as long-term funding can be more difficult to acquire in lower tiers of a research organisation (i.e., at departmental or individual research group level).
Third-party funding can mitigate this issue in the mid-term.
However, the participants considered central RSE groups more likely in the future,
due to the fact that there are already examples and activities from various universities in this direction
(examples include the universities of Jena, Heidelberg, and Göttingen, as well as several universities
in the UK where the concept of dedicated RSE groups began to emerge around 2013).
We expect funding in central groups to evolve from initially temporary positions to permanent positions,
as the groups prove themselves to be increasingly valuable and, hence, sustainable.
Embedded RSEs and central groups will need to network with each other,
and potentially request and coordinate support from external RSEs.
Within Germany at least, the distinction around whether these positions are classed as research staff or technical staff has so far created some discussion.
The authors hope, that 10 years from now this discussion has resolved towards a solution that enables the employment of RTPs more flexibly,
and thereby recognizing that the distinction between research staff and technical staff is too restrictive.

>>>>

\subsection{Impact of Alternative Career Paths for Researchers on the Structure of Research Institutes}
\begin{whatis}{}{career-paths}
As the RSE becomes establishedwe can see the emergence of more dedicated staff, i.e. other (digital) RTPs.
The legal requirements for employing people have become flexible enough to allow for the employment of more RTPs.
Work outside of the publication record is more readily accepted and enables researchers to build careers at the interface of academia, industry, and policy
(\eg\ in think tanks, NGOs, or organisations that do knowledge transfer).
\end{whatis}
RSEs can either be embedded in research groups (belonging to the mid-level university staff),
or be members of larger RSE units~\cite{Kempf2025-draft}.
The participants had the opinion that focusing on embedded RSEs would make onboarding new RSEs to existing projects easier.
This, however, comes with additional funding challenges,
as long-term funding can be more difficult to acquire in lower tiers of a research organisation (i.e., at departmental or individual research group level).
Third-party funding can mitigate this issue in the mid-term.
However, the participants considered central RSE groups more likely in the future,
due to the fact that there are already examples and activities from various universities in this direction
(examples include the universities of Jena, Heidelberg, and Göttingen, as well as several universities
in the UK where the concept of dedicated RSE groups began to emerge around 2013).
We expect funding in central groups to evolve from initially temporary positions to permanent positions,
as the groups prove themselves to be increasingly valuable and, hence, sustainable.
Embedded RSEs and central groups will need to network with each other,
and potentially request and coordinate support from external RSEs.
Within Germany at least, the distinction around whether these positions are classed as research staff or technical staff has so far created some discussion.
The authors hope that, 10 years from now, this discussion has resolved towards a solution that enables the employment of RTPs more flexibly,
and thereby recognizing that the distinction between research staff and technical staff is too restrictive.
